---
title: "Pruning Effects on Apple Tree Harvest"
author: Zack Nichols
date: "`r Sys.Date()`"
output: 
 github_document:
    toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# ABSTRACT

# BACKGROUND


```{r viremia, include=TRUE, eval=TRUE, warning=FALSE, fig.align='left', fig.width=4.75, fig.height=7}

```

# STUDY QUESTION and HYPOTHESIS

## Questions  
How does annual pruning influence the likelihood of achieving higher categorical harvest ratings among apple trees in Northern Utah orchards?

## Hypothesis 
If pruning enhances canopy light penetration and resource allocation within apple trees, then pruned trees will produce a significantly higher fruit average harvest rating compared to unpruned individuals.

## Prediction
If our hypothesis is correct, we expect to see that pruned trees have a higher proportion of *Above Average* harvest yield classifications, and a lower proportion of *Minimal* yields relative to unpruned trees, indicating a positive association between pruning and categorical harvest outcomes.

# METHODS


## Fill in 1st analysis e.g. barplots


```{r first-analysis, warning=FALSE}
install.packages("ggplot2") 
library("ggplot2")

pruning_data <- read.csv("pruning_data.csv")

ggplot(pruning_data, aes(x=pruningstatus, y=sizeofharvest)) + 
  geom_point(size=3, fill="gray50", alpha=.2) + # data point parameters
  geom_boxplot(fill= "lightgreen", alpha=0.6) + # box parameters
  xlab("Pruning Status of Tree") + # X-axis label
  ylab("Harvest Yield Grade") + # Y-axis label
  theme_classic() + # removes the grid lines in the background 
  theme(axis.text.x = element_text(size=15)) + # changes x axis text size
  theme(axis.title.x = element_text(size=20)) + # changes x axis title size
  theme(axis.title.x = element_text(face="bold")) + # changes x axis title to bold
  theme(axis.title.x = element_text(margin = margin(t = 10))) + # x/y axis distance 
  theme(axis.text.y = element_text(size=15)) + # changes y axis text size
  theme(axis.title.y = element_text(size=20)) + # changes y axis title size
  theme(axis.title.y = element_text(face="bold")) + # changes y axis title to bold
  theme(axis.title.y = element_text(margin = margin(r = 10))) # label distance

```
## Fill in 2nd analysis/plot e.g. generalized linear model

```{r}


```

# DISCUSSION

## Interpretation of 1st analysis (e.g. barplots)

## Interpretation of 2nd analysis (e.g. generalized linear model)

# CONCLUSION

# REFERENCES     

1. Komar N, Langevin S, Hinten S, Nemeth N, Edwards E, Hettler D, Davis B, Bowen R, Bunning M. Experimental infection of North American birds with the New York 1999 strain of West Nile virus. Emerg Infect Dis. 2003 Mar;9(3):311-22. https://doi.org/10.3201/eid0903.020628

2. ChatGPT. OpenAI, version Jan 2025. Used as a reference for functions such as plot() and to correct syntax errors. Accessed `r Sys.Date()`.