---
title: "Pruning Effects on Apple Tree Harvest"
author: Zack Nichols
date: "`r Sys.Date()`"
output: 
 github_document:
    toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# Libraries
library(ggplot2)
library(RColorBrewer)
library(plyr)

# Data Set
pruning_data <- read.csv(file="pruning_data.csv")


```

# ABSTRACT

# BACKGROUND

Successful apple production depends on a variety of factors, from environmental conditions to orchard management practices. One of these management practices, pruning, is recognized as one of the most essential techniques for maintaining tree productivity. Similar to tree species that do *not* produce fruit, apple trees naturally invest energy into vegetative growth, often producing dense canopies that limit internal light ability. This reduced light penetration can inhibit flower bud formation, restrict fruit development, and ultimately lower the overall harvest quality from the tree (Gomasta et al., 2024). Pruning is used to open the canopy, improve sunlight distribution across all leaves, and encourage the production of fruiting wood, making the practice a crucial component of orchard management

Beyond simply improving light exposure, pruning also influences other factors supplemental to fruit growth. These include air flow, disease pressure, plant stability, and growth distribution. Strategically removing crowded or shaded branches helps a tree to maintain vigor, while supporting consistent fruit production across seasons, and also increase the quality and development of the individual fruits. As a result, pruned trees are expected to achieve more reliable yields than unpruned plants, especially in the case of environments where environmental variability can compound stress on fruit development (Renquist, 2018).

In many commercial and research-oriented orchards, harvest success is recorded using categorical classifications when direct, quantitative counts of fruit number or biomass are unavailable. These categories provide a practical means of assessing overall harvest trends across large populations of trees, without comparing them to one another. Understanding whether pruning increases the likelihood of a higher harvest classification is important for optimizing orchard efficiency, participatory in regions like Northern Utah where unpredictable fluctuations in temperature, precipitations, and pollinator activity can affect fruit yield.

This study examines whether pruned apple trees are more likely to receive higher categorical harvest ratings compared to unpruned trees, providing insight into the role of pruning practices in determining yield outcomes.


```{r viremia, include=TRUE, eval=TRUE, warning=FALSE, fig.align='left', fig.width=4.75, fig.height=7}

```

# STUDY QUESTION and HYPOTHESIS

## Questions  
How does annual pruning influence the likelihood of achieving higher categorical harvest ratings among apple trees in Northern Utah orchards?

## Hypothesis 
If pruning enhances canopy light penetration and resource allocation within apple trees, then pruned trees will produce a significantly higher fruit average harvest rating compared to unpruned individuals.

## Prediction
If our hypothesis is correct, we expect to see that pruned trees have a higher proportion of *Above Average* harvest yield classifications, and a lower proportion of *Minimal* yields relative to unpruned trees, indicating a positive association between pruning and categorical harvest outcomes.

# METHODS


## 1st analysis


```{r first-analysis, warning=FALSE}
# one factor T test
# or ANOVA

```
## 2nd analysis

```{r, second-analysis, warning=T}
## Create a linear model determine whether pruning impacts apple harvest
apple_lm <- lm(harvest ~ pruned, data = apples)
resids <- apple_lm$residuals
qqnorm(resids)


```

## 3rd analysis

```{r, third-analysis, warning=T}
ggplot(pruning_data, aes(x=pruningstatus, y=sizeofharvest)) + 
  # point parameters
  geom_point(size=3, fill="gray50", alpha=.2) + # box point parameters
  geom_jitter(color="grey50", size=0.5, alpha=0.2) + # individual data points
  # box parameters
  geom_boxplot(aes(fill= pruningstatus), alpha=0.5, width = 0.6) + # box parameters 
  scale_fill_manual(values = c("Unpruned" = "forestgreen",
                               "Pruned" = "steelblue")) + 
  scale_color_manual(values = c("Unpruned" = "forestgreen",
                               "Pruned" = "steelblue")) + 
                      # makes unpruned/pruned different colors
  # Axis Labels + title
  xlab("Pruning Status of Tree") + # X-axis label
  ylab("Harvest Yield Grade") + # Y-axis label
  ggtitle("Harvest Yield Grade Across Pruning Practice") + 
  theme_classic() + # removes the grid lines in the background 
  theme(
    plot.title = element_text(hjust=0.5, size=22, face="bold"), # changes title format
    axis.text.x = element_text(size=15), # changes x axis text size
    axis.title.x = element_text(size=20, face="bold", margin = margin(t=10)), 
    axis.text.y = element_text(size=15), # changes y axis text size
    axis.title.y = element_text(size=20, face="bold", margin = margin(r = 10)), # changes y axis title size
    legend.position = "none"
  )

```
# DISCUSSION

## Interpretation of 1st analysis

## Interpretation of 2nd analysis 

# CONCLUSION

# REFERENCES     

1. Gomasta, J., Sarker, B., Haque, M., Anwari, A., Mondal, S., & Uddin, S. (2024). Pruning techniques affect flowering, fruiting, yield and fruit biochemical traits in guava under transitory sub-tropical conditions. Heliyon, e30064–e30064. https://doi.org/10.1016/j.heliyon.2024.e30064

2. Renquist, S. (2018, July 3). Fruit thinning. OSU Extension Service. https://extension.oregonstate.edu/catalog/em-9598-fruit-thinning

3. ChatGPT. OpenAI, version Jan 2025. Used as a reference for functions such as plot() and to correct syntax errors. Accessed `r Sys.Date()`.
